
%************************************************
\chapter{Problem Statement}
\label{sec:problem}
%************************************************

\section{Project Type}\label{sec:type}

In this bachelor thesis, I will be conducting attack path analyses on different internal vehicle network architectures.
I will then compare them based on which provides more security by introducing and using a tool that can be used to automate the evaluation of vehicular network security based on \gls{attack path} feasibility.
This bachelor thesis is done in cooperation with Mercedes-Benz Tech Innovation.\\

First, I will create multiple vehicular network architecture diagrams (\hyperref[sec:a1]{A1}).\\
Second, I will write a script-based tool that automizes the evaluation of said architectures
(\hyperref[sec:f1]{F1},
\hyperref[sec:f2]{F2},
\hyperref[sec:f3]{F3},
\hyperref[sec:f4]{F4},
\hyperref[sec:f5]{F5}).\\
Every architecture consists of \gls{ecu}s (nodes) and bus systems (edges) connecting the ECUs, as well as interfaces, each of which has an attack feasibility rating.
The feasibilities are scaled based on information provided by security experts.\\

Next, I will decide on a criteria (\hyperref[sec:a2]{A2}), how to rate the different topologies.
To decide on a criteria, I will conduct a survey with security architects and pentesters at MBTI, in which they will rate the different architectures based on their experience and knowledge.\\
The survey will be used to calibrate my own criteria.\\

Finally, I will compare the architectures to conclude which architecture offers the most security (\hyperref[sec:a3]{A3}).\\

The thesis will be both exploratory and implementory in nature, with a focus on the latter.


\section{Thesis Questions}\label{sec:thesis-questions}

The questions this thesis will answer include:

\begin{itemize}
    \item How can different E/E architectures be rated based on attack paths?
    \item How secure is the given vehicular network architecture?
    \item What architectural approach makes a network safer than others?
    \begin{itemize}
        \item How do small changes in network positioning affect the network security overall?
        \item How do simple and more branched out networks compare in terms of security?
    \end{itemize}
\end{itemize}



