%************************************************
\chapter{Requirements Analysis}
\label{chp:requirements}
%************************************************

\graffito{\begin{shaded}\textbf{just implementation and constructive projects}, remove this chapter if your project is of any other type.\end{shaded}}

\begin{shaded}
\noindent
goal of this project, approach

\medskip
\noindent
give a detailed definition of the initial problem that requires a solution, 
respectively the challenges of transferring or synthesizing a concept into a construction or realization

\medskip
\noindent
define use cases and the deduced requirements for the implementation or construction
\end{shaded}

In this thesis you have to make attack path analyses on different internal vehicle network architectures and compare them
based on which provides more security with regards attack
paths.\\

The first step would be creating multiple different architecture diagrams. Then you have to write a program, which reads
files of a vehicle network topology, maps this to a list of entry point and target ECUs, and generates a list of all possible
attack paths. To get a quick and early result, this list should
be sorted by the number of hops over each gateway. The next
step would be giving each entry point, gateway and connection a rating on how big the attack feasibility for this element
is. Then, attack paths can be calculated - e.g. with the formula
of the paper ”ThreatSurf A method for automated Threat Surface assessment”. At last, you have to decide on a criteria on
how to rate the different topologies and compare them with
it.


%************************************************
\chapter{Functional Specification}
\label{chp:specification}
%************************************************

\graffito{\begin{shaded}\textbf{just implementation and constructive projects}, remove this chapter if your project is of any other type.\end{shaded}}

\begin{shaded}
\noindent
specification of technical requirements, entities, interfaces, protocols and procedures that will be employed to meet the identified requirements,

\medskip
\noindent
interdependencies of technical, human or environmental entities
\end{shaded}
