\chapter{Introduction}
\label{chp:introduction}

\section{Research field: Cybersecurity of vehicular networks}
\label{sec:field}

Modern vehicles are becoming increasingly reliant on technology, with a wide range of systems and components being connected to the internet and each other. 
This includes everything from infotainment systems and navigation to advanced driver assistance systems and autonomous driving features.
ISO 26262 describes these so-called "\gls{e/e} systems" as systems that consist of electrical and electronic elements and components.
Examples include \gls{ecu}s, sensors, actuators, connections, and communication systems like \gls{can}, Ethernet, and Bluetooth \cite{iso26262}.
As these technologies become more and more important, strong cybersecurity measures are also becoming increasingly important. 
Cybercriminals are constantly finding new ways to exploit vulnerabilities in these systems which can have serious consequences for users, 
which includes their safety, privacy, finances, and operational viability \cite{iso21434}. 
Ensuring the safety and security of connected vehicles is crucial to the success of the emerging world of connected and autonomous transportation.
\\

The internal vehicular network architecture plays a crucial role in the overall cybersecurity of a modern vehicle as it determines how different vehicle systems and components are connected and communicate.
Attack paths play an important role in vehicle networks and security because they define the specific routes that a malicious actor might use to attack a vehicle's systems or networks. 
A well-designed internal vehicular network architecture can help minimize cyberattack risk.
With the increasing number of systems and components that are connected to the internet and each other, the complexity of the internal vehicular network architecture is also increased. 
This can make it more challenging to design and implement effective cybersecurity measures as more potential points of vulnerability arise that need to be addressed. 
Therefore, companies developing connected and autonomous vehicles need to prioritize cybersecurity in designing their internal vehicular network architecture, 
which can help to ensure the safety and security of the vehicle, as well as protect the privacy and personal data of drivers and passengers.
\\

Security testing is, therefore, a crucial part of the development process.
A \gls{tara}, for example, is performed early during the development process to identify and prioritize potential risks.
This information can be used to guide the design and implementation of controls or countermeasures to reduce or mitigate these risks.
Once the system is developed, e.g. a penetration test is carried out to validate the effectiveness of these controls and to identify any remaining vulnerabilities. 
The results of the pentest can then be incorporated back into the \gls{tara} process to update the risk assessment and prioritize future risk mitigation efforts. 
For example, companies can implement appropriate security measures by understanding the attack paths that might be used to gain unauthorized access to a vehicle's systems. 
These include encryption, authentication protocols, and firewall systems, etc. to protect against cyber threats.\\
In this way, the combination of a pentest and \gls{tara} provides a complete and iterative approach to security assessment.\\
However, security testing is often carried out in the late stages of development, which can lead to the discovery of vulnerabilities at a time when it is more difficult and costly to address.
Additionally, they are considered to be a skill-based activity that is still carried out manually.
It requires a high level of expertise and experience with other cybersecurity tools and techniques. 
The increased complexity of modern vehicles and the arduous nature of state-of-the-art security testing methods make it more unfeasible for companies to conduct them as is done now.\\
Thus, the need for a more efficient approach to improve overall security testing is evident. 
By automizing the process of evaluating automotive network architectures and their attack paths, 
potential vulnerabilities can be assessed early on in the development process resulting in overall more efficient security testing, improving the flexibility, costs and accelerateing it.\\

In this thesis we will explore a \nameref{chp:criteria} that can be used to evaluate the security of vehicular network architectures and how it can be used to automate the process of security testing.\\
This criteria is based on the experience of security experts and the results of a \nameref{sec:survey} conducted by the author of this thesis.
This survey will feature a training set of architectures which the security experts will evaluate and give feedback on the architectures' security.
Their ratings and feedback will be used as calibration for the criteria.\par
In order to automize the evaluation of vehicular network architectures, we will use a graph-based approach to model the network architecture and its attack paths
and develop a \nameref{chp:tool} that can be used to evaluate the security of vehicular network architectures based on the criteria in an automated manner.
The \nameref{chp:compeval} contains another set of architectures that will be used to test the criteria and the tool.
In the end, we will evaluate the results of the tool and the criteria and discuss the limitations and potential improvements.

\section{Thesis Questions}
\label{sec:thesis-questions}

The questions this thesis will answer include:

\begin{itemize}
    \item How can different E/E architectures be rated based on attack paths?
    \item How secure is the given vehicular network architecture?
    \item What architectural approach makes a network safer than others?
    \begin{itemize}
        \item How do small changes in network positioning affect the network security overall?
        \item How do simple and more branched out networks compare in terms of security?
    \end{itemize}
\end{itemize}